### 2018-11-07 Docs/architecture/architecture.tex

\maestro\ can use adaptive mesh refinement to focus resolution on
complex regions of flow.  For Cartesian/plane-parallel geometries, all
cells at the same height must be at the same level of refinement.
This restriction is to allow for the base state to directly align with
the Cartesian state everywhere.  For spherical geometries, there is no
such restriction (again, see Figure~\ref{fig:base_state}).
The \maestro\ grids are managed by the \boxlib\ library, which is
distributed separately.

@@

\begin{figure}[]
\centering
\includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
\caption[\maestro\ `ecosystem'] {\label{fig:arch:eco} The basic
  \maestro\ `ecosystem'.  Here we see the different packages that
  contribute to building the {\tt reacting\_bubble} problem in \maestro.  The
  red directories are part of most standard \maestro\ build.  The
  purple lines show the directories that are pulled in through
  various Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).}
\end{figure}

Building \maestro\ requires both the \maestro-specific source
files (distributed in the {\tt MAESTRO/} directory), and the
\boxlib\ library (distributed separately, consisting of the {\tt BoxLib/} directory).
\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
only uses the Fortran portions of \boxlib.  Figure~\ref{fig:arch:eco}
shows the relationship between the different packages, highlighting
what goes into defining a specific \maestro\ problem.

@@

  \item {\tt Microphysics/}\footnote{Note: many more \compatible\ routines are available in the separate \microphysics\ git repo}

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
\caption[\boxlib\ directory structure] {\label{fig:arch:boxlib} The
  basic \boxlib\ directory structure.  The directories used by
  \maestro\ are indicated in red.}
\end{figure}

The \boxlib\ directory structure is shown in
Figure~\ref{fig:arch:boxlib}.  The subset of the directories that are
used by \maestro\ are:
\begin{itemize}
\item {Src/}

  The main \boxlib\ source directory.  In \maestro, we only use the
  Fortran source files.  The core directories are:

@@

    The Fortran \boxlib\ files.  This is a library for describing
    meshes consisting of a union of boxes.  The \boxlib\ modules
    define the basic datatypes used in \maestro.  \boxlib\ also
    provides the routines that handle the parallelization and I/O.

@@

    The \boxlib\ linear solvers---these are used to solve elliptic
    problems in the \maestro\ algorithm.

@@

  Various tools used for building \boxlib\ applications.  Here we use:

@@

Finally the {\tt BoxLib/Tools/Postprocessing/F\_Src} package provides simple
Fortran-based analysis routines (e.g.\ extract a line from a
multidimensional dataset) that operate on BoxLib datasets.  These are
described in \S~\ref{sec:analysis}.  Several sub-directories with
python-based routines are also here.  These are described both in
\S~\ref{sec:analysis} and \S~\ref{sec:vis:python}.

@@

\section{BoxLib Data Structures}

\maestro's gridding is handled by the BoxLib library, which
contains the most fundamental objects used to construct parallel
block-structured AMR applications---different
regions of the domain can have different spatial resolutions.
At each level of refinement, the region covered by that level is divided
into grids, or boxes.  The entire computational domain is covered by
the coarsest (base) level of refinement, often called level $\ell=0$, either by one
grid or divided into many grids.
Higher levels of refinement have cells that are finer by a ``refinement ratio''
(typically 2).  The grids are properly nested in the sense that the union 
of grids at level $\ell+1$ is contained in the union of grids at level $\ell$.
Furthermore, the containment is strict in the sense that, except at physical 
boundaries, the level $\ell$ grids are large enough to guarantee that there is
a border at least $n_{\rm buffer}$ level $\ell$ cells wide surrounding each level
$\ell +1$ grid (grids at all levels are allowed to extend to the physical
boundaries so the proper nesting is not strict there).  
For parallel computations, the boxes are spread across processors, in
a fashion designed to put roughly equal amounts of work on each
processor (load balancing).
\MarginPar{ghost cells?}

\begin{figure}[t]
\centering
\includegraphics[width=6.5in]{\archfigpath/data_loc2}
\caption[Data-centerings on the grid]
  {\label{fig:dataloc} Some of the different data-centerings:
  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
  both the $x$- and $y$-directions.  Note that for nodal data, the
  integer index corresponds to the lower boundary in that direction.
  In each of these centerings, the red point has the same indices:\ (1,2).
  Not shown is the case where data is nodal in the $y$-direction only.}
\end{figure}
On a grid, the data can be stored at cell-centers, on a face/edge, or
on the corners.  In BoxLib, data that is on an edge is termed `nodal'
in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
corners is nodal in all spatial directions.  In \maestro, the state
data (density, enthalpy, velocity, $\ldots$) is generally
cell-centered.  Fluxes are nodal in the direction they represent.
A few quantities are nodal in all directions (e.g.\ $\phi$ used in
the final velocity projection).

To simplify the description of the underlying AMR grid, BoxLib
provides a number of Fortran types.  We briefly summarize the major
data types below.  A more extensive introduction to \boxlib\ is 
provided by the \boxlib\ User's Guide, distributed with the library.

@@

\begin{figure}[t]
\centering
\includegraphics[width=4.0in]{\archfigpath/index_grid2}
\caption[Single-level grid structure]
{\label{fig:boxes} Three boxes that comprise a single level.  At this
  resolution, the domain is 20$\times$18 zones.  Note that the
  indexing in BoxLib starts with $0$.}
\end{figure}

@@

The procedure for accessing the data in each grid managed by the
\multifab\ is shown in \S~\ref{sec:example}.  Subroutines to add,
multiply, or divide two \multifab s exist, as do subroutines to copy
from one \multifab\ to another---see
\code{BoxLib/Src/F\_BaseLib/multifab.f90} for the full list of
routines that work with \multifab s.

@@

Presently there is a single EOS that comes with \maestro, {tt gamma\_law\_general},
but many more are available through the external \microphysics\ repo\footnote{\microphysics\ is
available at \url{https://github.com/BoxLib-Codes/Microphysics}.  \maestro\ will
find it via the {\tt MICROPHYSICS\_HOME} environment variable}.  The \microphysics\
EOSs share the same interface and can be compiled into \maestro\ directly.  
Here are the more popular EOSs:
\begin{itemize}
\item {\tt helmholtz} represents a general stellar equation 
      of state, consisting of nuclei (as an ideal gas), radiation,
      and electrons (with arbitrary degeneracy and degree of relativity).
      This equation of state is that described in \cite{timmes_eos}.

@@

\section{BoxLib Helper Modules}

There are a large number of modules in {\tt boxlib/} that provide
the core functionality for managing grids.  Here we describe
the most popular such modules.

@@

\noindent Here, the {\tt bl\_types} and \code{multifab\_module}
modules bring in the basic BoxLib data types. Specifically, here,
\code{bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
declaring double precision data, and \code{multifab\_module} defines
the \multifab\ data type.  The \code{ml\_layout\_module} defines the
datatype for a \mllayout---many routines will take an \mllayout to
allow us to fill ghostcells.  The \code{variables} module is a
\maestro\ module that provides integer keys for indexing the state
arrays.  In this case the integer {\tt rho\_comp} refers to the
location in the state array corresponding to density.

@@

You generally won't see calls in the MAESTRO source code to these subroutines,
as there is now a special BoxLib subroutine, {\tt ml\_restrict\_and\_fill},
that takes an array of multifabs at different level, and in order calls:
(1) {\tt ml\_cc\_restriction\_c}, (2) {\tt multifab\_fill\_boundary},
(3) {\tt multifab\_physbc}, and (4) {\tt multifab\_fill\_ghost\_cells}.
These four subroutines are called in such a way to avoid extra
ghostcell filling, saving on communication time.  You can specify the
starting component, starting boundary condition component, 
the number of components, the number of ghost cells,
and whether or not you want to use the same boundary condition component
for all variables.

@@

When \maestro\ is run, the boundary condition parameters are read in
from the input file and used to build the \bctower.  The
\bctower\ consists of a \bclevel\ object for each level of resolution
in the simulation.  The \bclevel\ contains 3 different descriptions of
the boundary conditions for each box in the domain at that level of
refinement: {\tt phys\_bc\_level\_array}, {\tt adv\_bc\_level\_array},
and {\tt ell\_bc\_level\_array}.  In all cases, the boundary
conditions are specified via integer values that are defined in {\tt
  bc\_module} (part of \boxlib).

@@

There is an extensive regression test suite for \boxlib\ that works with
\maestro.  Full details, and a sample \maestro\ configuration file are
provided in the \boxlib\ User's Guide and source.

>>>>

\maestro\ can use adaptive mesh refinement to focus resolution on
complex regions of flow.  For Cartesian/plane-parallel geometries, all
cells at the same height must be at the same level of refinement.
This restriction is to allow for the base state to directly align with
the Cartesian state everywhere.  For spherical geometries, there is no
such restriction (again, see Figure~\ref{fig:base_state}).
The \maestro\ grids are managed by the \fboxlib\ library\footnote{This is a pure Fortran library that 
has a lot of similarities to the \cpp\ \amrex\ library}, which is
distributed separately.

@@

\begin{figure}[]
\centering
\includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
\caption[\maestro\ `ecosystem'] {\label{fig:arch:eco} The basic
  \maestro\ `ecosystem'.  Here we see the different packages that
  contribute to building the {\tt reacting\_bubble} problem in \maestro.  The
  red directories are part of most standard \maestro\ build.  The
  purple lines show the directories that are pulled in through
  various Makefile variables ({\tt FBOXLIB\_HOME}, {\tt NETWORK\_DIR},
  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).}
\end{figure}

Building \maestro\ requires both the \maestro-specific source
files (distributed in the {\tt MAESTRO/} directory), and the
\fboxlib\ library (distributed separately, consisting of the {\tt FBoxLib/} directory).
Figure~\ref{fig:arch:eco}
shows the relationship between the different packages, highlighting
what goes into defining a specific \maestro\ problem.

@@

  \item {\tt Microphysics/}\footnote{Note: many more compatible routines are available in the separate \microphysics\ git repo}

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.6]{\archfigpath/amrex_directory2}
\caption[\amrex\ directory structure] {\label{fig:arch:amrex} The
  basic \amrex\ directory structure.  The directories used by
  \maestro\ are indicated in red.}
\end{figure}

The \amrex\ directory structure is shown in
Figure~\ref{fig:arch:amrex}.  The subset of the directories that are
used by \maestro\ are:
\begin{itemize}
\item {Src/}

  The main \amrex\ source directory.  In \maestro, we only use the
  Fortran source files.  The core directories are:

@@

    The Fortran \amrex\ files.  This is a library for describing
    meshes consisting of a union of boxes.  The \amrex\ modules
    define the basic datatypes used in \maestro.  \amrex\ also
    provides the routines that handle the parallelization and I/O.

@@

    The \amrex\ linear solvers---these are used to solve elliptic
    problems in the \maestro\ algorithm.

@@

  Various tools used for building \amrex\ applications.  Here we use:

@@

Finally the {\tt FBoxLib/Tools/Postprocessing/F\_Src} package provides simple
Fortran-based analysis routines (e.g.\ extract a line from a
multidimensional dataset) that operate on \amrex\ datasets.  These are
described in \S~\ref{sec:analysis}.  Several sub-directories with
python-based routines are also here.  These are described both in
\S~\ref{sec:analysis} and \S~\ref{sec:vis:python}.

@@

\section{\fboxlib\ Data Structures}

\maestro's gridding is handled by the \fboxlib\ library, which
contains the most fundamental objects used to construct parallel
block-structured AMR applications---different
regions of the domain can have different spatial resolutions.
At each level of refinement, the region covered by that level is divided
into grids, or boxes.  The entire computational domain is covered by
the coarsest (base) level of refinement, often called level $\ell=0$, either by one
grid or divided into many grids.
Higher levels of refinement have cells that are finer by a ``refinement ratio''
(typically 2).  The grids are properly nested in the sense that the union 
of grids at level $\ell+1$ is contained in the union of grids at level $\ell$.
Furthermore, the containment is strict in the sense that, except at physical 
boundaries, the level $\ell$ grids are large enough to guarantee that there is
a border at least $n_{\rm buffer}$ level $\ell$ cells wide surrounding each level
$\ell +1$ grid (grids at all levels are allowed to extend to the physical
boundaries so the proper nesting is not strict there).  
For parallel computations, the boxes are spread across processors, in
a fashion designed to put roughly equal amounts of work on each
processor (load balancing).
\MarginPar{ghost cells?}

\begin{figure}[t]
\centering
\includegraphics[width=6.5in]{\archfigpath/data_loc2}
\caption[Data-centerings on the grid]
  {\label{fig:dataloc} Some of the different data-centerings:
  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
  both the $x$- and $y$-directions.  Note that for nodal data, the
  integer index corresponds to the lower boundary in that direction.
  In each of these centerings, the red point has the same indices:\ (1,2).
  Not shown is the case where data is nodal in the $y$-direction only.}
\end{figure}
On a grid, the data can be stored at cell-centers, on a face/edge, or
on the corners.  In \fboxlib, data that is on an edge is termed `nodal'
in that direction (see Figure~\ref{fig:dataloc}).  Data that is on the
corners is nodal in all spatial directions.  In \maestro, the state
data (density, enthalpy, velocity, $\ldots$) is generally
cell-centered.  Fluxes are nodal in the direction they represent.
A few quantities are nodal in all directions (e.g.\ $\phi$ used in
the final velocity projection).

To simplify the description of the underlying AMR grid, \fboxlib\
provides a number of Fortran types.  We briefly summarize the major
data types below.  A more extensive introduction to \fboxlib\ is 
provided by the \fboxlib\ User's Guide, distributed with the library.

@@

\begin{figure}[t]
\centering
\includegraphics[width=4.0in]{\archfigpath/index_grid2}
\caption[Single-level grid structure]
{\label{fig:boxes} Three boxes that comprise a single level.  At this
  resolution, the domain is 20$\times$18 zones.  Note that the
  indexing in \fboxlib\ starts with $0$.}
\end{figure}

@@

The procedure for accessing the data in each grid managed by the
\multifab\ is shown in \S~\ref{sec:example}.  Subroutines to add,
multiply, or divide two \multifab s exist, as do subroutines to copy
from one \multifab\ to another---see
\code{FBoxLib/Src/F\_BaseLib/multifab.f90} for the full list of
routines that work with \multifab s.

@@

Presently there is a single EOS that comes with \maestro, {tt gamma\_law\_general},
but many more are available through the external \microphysics\ repo\footnote{\microphysics\ is
available at \url{https://github.com/starkiller-astro/Microphysics}.  \maestro\ will
find it via the {\tt MICROPHYSICS\_HOME} environment variable}.  The \microphysics\
EOSs share the same interface and can be compiled into \maestro\ directly.  
Here are the more popular EOSs:
\begin{itemize}
\item {\tt helmholtz} represents a general stellar equation 
      of state, consisting of nuclei (as an ideal gas), radiation,
      and electrons (with arbitrary degeneracy and degree of relativity).
      This equation of state is that described in \cite{timmes_eos}.

@@

\section{\fboxlib\ Helper Modules}

There are a large number of modules in {\tt FBoxLib/} that provide
the core functionality for managing grids.  Here we describe
the most popular such modules.

@@

\noindent Here, the {\tt bl\_types} and \code{multifab\_module}
modules bring in the basic \fboxlib\ data types. Specifically, here,
\code{bl\_types} defines {\tt dp\_t} which is the {\tt kind} used for
declaring double precision data, and \code{multifab\_module} defines
the \multifab\ data type.  The \code{ml\_layout\_module} defines the
datatype for a \mllayout---many routines will take an \mllayout to
allow us to fill ghostcells.  The \code{variables} module is a
\maestro\ module that provides integer keys for indexing the state
arrays.  In this case the integer {\tt rho\_comp} refers to the
location in the state array corresponding to density.

@@

You generally won't see calls in the MAESTRO source code to these subroutines,
as there is now a special \fboxlib\ subroutine, {\tt ml\_restrict\_and\_fill},
that takes an array of multifabs at different level, and in order calls:
(1) {\tt ml\_cc\_restriction\_c}, (2) {\tt multifab\_fill\_boundary},
(3) {\tt multifab\_physbc}, and (4) {\tt multifab\_fill\_ghost\_cells}.
These four subroutines are called in such a way to avoid extra
ghostcell filling, saving on communication time.  You can specify the
starting component, starting boundary condition component, 
the number of components, the number of ghost cells,
and whether or not you want to use the same boundary condition component
for all variables.

@@

When \maestro\ is run, the boundary condition parameters are read in
from the input file and used to build the \bctower.  The
\bctower\ consists of a \bclevel\ object for each level of resolution
in the simulation.  The \bclevel\ contains 3 different descriptions of
the boundary conditions for each box in the domain at that level of
refinement: {\tt phys\_bc\_level\_array}, {\tt adv\_bc\_level\_array},
and {\tt ell\_bc\_level\_array}.  In all cases, the boundary
conditions are specified via integer values that are defined in {\tt
  bc\_module} (part of \fboxlib).

@@

There is an extensive regression test suite for \amrex\ that works with
\maestro.  Full details, and a sample \maestro\ configuration file are
provided in the \amrex\ User's Guide and source.
